Biogeography and Environmental Conditions Shape Phage and Bacteria Interaction Networks Across the Healthy Human Microbiome

### 2017-05-05 doc/manuscript.Rmd
Went back and moved changed from md to rmd. I need to make it clearer that md is an int file

# Abstract
Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease. Most previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities. Such approaches are not able to capture how these communities interact to maintain stability and share information (i.e. horizontal gene transfer). We developed and implemented a network-based analytical approach to provide an initial understanding of phage-bacteria network diversity throughout the human body. We accomplished this by building a machine learning algorithm to predict which phages infected which bacteria in a given microbiome. This algorithm was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts. We organized the predicted interactions into networks that allowed us to evaluate phage-bacteria connectedness across the human body. We found that gut and skin network structures were person-specific, and were not conserved among cohabitating family members. High-fat diets and obesity were associated with less connected networks. There were significant differences in network structure between skin sites, with those exposed to the external environment being less connected and more prone to instability. Together this study characterized the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics. This work provides a baseline for future studies to investigate the role of ecological networks in conditions such as disease states.

@@

Viruses and bacteria are critical components to the human microbiome and play important roles in health and disease. Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb]. Altered viromes (virus communities consisting primarily of bacteriophages in humans) have also been associated with various diseases and environmental perturbations including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj]. These virus communities act in concert with bacteria as one single overall community [@Haerter:2014ii]. Individual phages within these communities are capable of lysing their bacterial hosts, modulating their functionality through processes such as horizontal gene transfer (i.e. transduction) [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn], and altering bacterial host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja].

Most previous work has relied on studying bacterial and phage communities in isolation, reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj]. In reality, bacteria and phage communities are dynamic and complex. They frequently share genetic information and work together to maintain stable ecosystems. Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev]. The common approach of studying phage and bacterial communities separately has been unable to capture how these communities interact and influence one another; however relationship-based network approaches allow us to capture this information. We aimed to bridge this knowledge gap by leveraging machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships. We focused on characterizing human skin and gut virome-microbiome network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.

We investigated human microbiome-virome networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei]. Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc]. This metagenomic interaction inference model was powered beyond previous models by its inclusion of protein interaction data as well as both negative and positive interactions [@Edwards:2015iz, @Roux:2016cc]. Through this approach, we were able to go beyond contemporary isolated methods to provide a basic understanding of the network dynamics associated with healthy human phage and bacteria communities. By building and utilizing a microbiome-virome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability. Through an improved understanding of the healthy state of network structures across the human body, we will empower future studies to begin investigating how these community structures change in disease and their overall impact on human health.

@@

## Cohort Curation and Sample Processing
We studied the differences in virome-microbiome interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences. Our study contained three datasets that studied impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei]. These datasets were selected because they were subjected to virus-like particle (VLP) purification to eliminate other organismal DNA and allow for direct assessment of both the extracellular and actively replicating intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].

The bacterial and viral sequences from these publications were quality filtered and assembled into contigs. We generated clusters of related bacterial and phage contigs by k-mer frequency and co-abundance using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**. Contig lengths within clusters ranged between $10_{3}$ and $10_{5.5}$, confirming the majority of contigs did not represent whole genomes alone **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**. We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs). These OGUs are conceptually similar to the operational taxonomic unit (OTU) and operational protein family (OPF) definitions used for grouping highly similar 16S rRNA gene and open reading frame sequences, respectively [@Schloss:2008ht].

@@

Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) targeting of phages by bacterial CRISPR spacer sequences, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq]. The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**. The most important predictor in the model was amino acid similarity between genes (blastx), followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**. Protein family interactions were moderately important to the model, and CRISPRs were not informative **(Figure \ref{RocCurve} B)**. Approximately one third of the training set relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.

@@

In addition to diet, we also observed an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**. The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)** but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**. These results suggest that the obesity network was overall less connected, having microbes further from all other microbes within the community.

## Individuality of Microbial Networks
Skin and gut community membership and diversity are highly personal, with people remaining more similar to themselves than to other people over time [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih]. We therefore hypothesized that this personal conservation extended to microbiome network structure. we addressed this hypothesis by calculating the degree of dissimilarity between each subject's network, based on phage and bacteria abundance as well as centrality. We quantified phage and bacteria centrality within each sample graph using the weighted eigenvector centrality metric. This metric defines central phages as those that are highly abundant and infect many distinct bacteria which themselves are abundant and infected by many other phages. Similarly, bacterial centrality was defined as those bacteria that were both abundant and connected to numerous phages that were themselves connected to many bacteria. We then calculated the similarity of community networks using the weighted eigenvector centrality of all nodes between all samples. Samples with similar network structures were interpreted as having similar capacities for maintaining stability and transmitting genetic material.

@@

Skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**) and more adequately confirmed by our larger skin dataset. This distribution was similar when separated by anatomical sites, with most sites also being statistically significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.

## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
Extensive work has illustrated differences in the healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz]. These communities vary by degree of skin moisture, oil, and environmental exposure. We hypothesized that like microbial composition and diversity, microbial network structure would be specific to anatomical sites. We tested this hypothesis by evaluating the changes in network structure between anatomical sites within our skin dataset.

We quantified the average centrality of each sample using the weighted eigenvector centrality metric. We found that intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**. We also found that skin sites that were protected from the environment (occluded) were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.

We supplemented this analysis by comparing the network signatures using the centrality dissimilarity approach described above. The dissimilarity between samples was a function of shared relationships, degree of centrality, and bacteria/phage abundance. When using this supplementary approach, we found that network structures significantly clustered by moisture, sebaceous, and intermittently moist status **(Figure \ref{skinnetwork} C,E)**. We also found that occluded sites were significantly different from exposed and intermittently occluded sites, but there was no difference between exposed and intermittently occluded sites **(Figure \ref{skinnetwork} D,F)**. These findings provide further support that skin microbiome network structure differs significantly between skin sites.

# Discussion
Foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez]. We similarily aimed to provide an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how microbiome networks differ in disease states. We developed and implemented a network-based analytical approach that we used to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity. Through the use of network theory, we leveraged extensive analytical opportunities that could be applied to understand complex ecological communities. We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact (e.g. horizontal gene transfer, modulated bacterial gene expression, alterations in abundance). By pursuing this goal, we aimed to provide an initial understanding of human microbiome networks that will power future studies to use similar approaches in other settings, such as disease states.

Just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were also conserved within individuals over time. Gut network structure was not conserved among family members. These findings suggested that microbiome network properties, such as stability and the potential for horizontal gene transfer, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions such as climate and diet. The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.

@@

In addition to diet, the skin microbiome network structure varied by skin environment. Network structure differed between environmentally exposed and occluded skin sites. The sites under greater environmental fluctuation and exposure, the exposed and intermittently exposed sites, were less connected and therefore were predicted to have a higher propensity for instability. Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites. Togteher this data suggested body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.

Our results represent a step toward understanding the microbiome through interspecies relationships. Although these findings are informative, there are certainly caveats that should be noted. First, although our infection classification model is advantageous over existing models, we recognize that, like most classification models, ours is only as good as its training set. Large scale experimental screens for phage and bacteria infectious interactions will provide more robust model training and therefore improved model performance. Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve performance. Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago. These archived datasets were limited by the technology and costs of the time, meaning the datasets were poorly powered for the statistical analysis we strive for today. Despite these limitations, we were able to present initial conclusions; however, follow-up studies will be required to validate our findings. These methods and results will be important for informing design and interpretation of future studies and for highlighting the importance of bacteria-virus interaction profiles as a characteristic of human-associated microbial communities.

The virome-microbiome relationship structure, and therefore the potential for microbiome genetic transfer and stability maintenance, differed significantly between body sites. Other environmental factors, such as diet and individuality, also influenced relationship structures of the microbiome. This information builds on previous work by suggesting that bacteria and phages not only preferentially colonize different body sites, but also interact differently, allowing for different communication structures and capacities for maintaining stability. Our results provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.

>>>>

# Abstract
Viruses and bacteria are critical components of the human microbiome and play important roles in health and disease. Most previous work has relied on studying bacteria and phages in isolation, reducing them to two separate communities. Such approaches are not able to capture how these communities interact to maintain stability and share information (e.g., horizontal gene transfer <!-- as this isn't the only way to share information, use eg instaed of ie. Also, is there work that demonstrates how these communities interact to maintain stability? By stability, do we mean maintain diversity? What are the other outcomes of interactions? e.g., (antagonistic) co-evolution, etc. I know these aren't elements to fit into an Abstract, so staying tuned as to whether they appear later, as well-->). We developed and implemented a network-based analytical approach to describe phage-bacteria network diversity throughout the human body. We accomplished this by building a machine learning algorithm to predict which phages infected which bacteria in a given microbiome. This algorithm was applied to paired viral and bacterial metagenomic sequence sets from three previously published human cohorts. We organized the predicted interactions into networks that allowed us to evaluate phage-bacteria connectedness across the human body. We found that gut and skin network structures were person-specific and were not conserved among cohabitating family members. High-fat diets and obesity were associated with less connected networks. There were significant differences in network structure between skin sites, with those exposed to the external environment being less connected and more prone to instability. This study quantified and contrasted the diversity of virome-microbiome networks across the human body and illustrated how environmental factors may influence phage-bacteria interactive dynamics. This work provides a baseline for future studies to investigate the role of virus-microbe ecological networks in advancing our understanding of conditions, such as disease states.

@@

Viruses and bacteria are critical components to the human microbiome and play important roles in health and disease. Bacterial communities have been associated with diseases including a wide range of skin conditions [@Hannigan:2013im], acute and chronic wound healing conditions [@Hannigan:2014be; @Loesche:2016ev], and gastrointestinal diseases including inflammatory bowel disease [@He:2016ch; @Norman:2015kb], *Clostridium difficile* infections [@Seekatz:2016fz], and colorectal cancer [@Zackular:2014fba; @Baxter:2014hb]. Altered viromes (virus communities consisting primarily of bacteriophages) have also been associated with various diseases and perturbations, including inflammatory bowel disease [@Norman:2015kb; @Manrique:2016dx], periodontal disease [@Ly:2014ew], spread of antibiotic resistance [@Modi:2013fia], and others [@Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj]. Viruses interact with bacteria and archaea and together they comprise one total microbial community [@Haerter:2014ii]. Viruses influence the living microbes of these communities by, e.g., lysing their microbial hosts, modulating their functionality through altered host gene expression [@Lindell:2005gz, @Tyler:2013fl, @Hargreaves:2014ja], influencing evolutionary processes, e.g., through horizontal gene transfer  [@Moon:2015fa, @Modi:2013fi, @Ogg:1981th, @Frost:2005dn] or co-evolution <!-- Koskella, B. & Brockhurst, M.A., 2014, Bacteria-phage coevolution as a driver of ecological and evolutionary processes in microbial communities, FEMS microbiology reviews    IMPORTANT, check out the abstract and intro of this paper--I think well-reviewed here. Bringing in a few of these examples to demistify what we are suggesting our network edges to imply will benefit this work and help reviewers appreciate the relevance. -->, and by altering ecosystem processes and elemental stoichiometry <!-- Jover, L.F., Effler, T.C., Buchan, A., Wilhelm, S.W. & Weitz, J.S., 2014, The elemental composition of virus particles: implications for marine biogeochemical cycles, Nature reviews. Microbiology, 12(7), pp. 519-28 -->.

Most previous work has relied on studying bacterial <!-- before I go through and change all instances of bacteria to something that might include archaea too, at least where shotgun metagenomics is used, I'll pose "is this intentional?" If not, I suggest changing it --> and phage communities in isolation, reducing them to two separate communities [@Norman:2015kb; @Manrique:2016dx; @Ly:2014ew; @Monaco:2016ita; @Hannigan:2015fz; @Minot:2011ez; @SantiagoRodriguez:2015gd; @Abeles:2015dy; @Abeles:2014kj]. In reality, bacteria and phage communities are dynamic and complex. They frequently share genetic information and work together to maintain stable ecosystems. Removal of bacteria or phage can disrupt or even collapse those ecosystems [@Haerter:2014ii; @Harcombe:2005fd; @Middelboe:2001fl; @Poisot:2011jc; @Thompson:2012ki; @Moebus:1981kp; @Flores:2013hc; @Poisot:2012fh; @Poisot071993; @Flores:2011bh; @Jover:2015ev]. Relationship-based network approaches allow us to capture this information. We leveraged machine learning and graph theory techniques to characterize human bacterial and phage communities by their relationships. We focused on characterizing human skin and gut virome-microbiome network diversity to provide a foundation for further studies of disease network dynamics and to gain broader insights into human microbiome network diversity across different body sites.

We investigated human microbiome-virome networks by utilizing three published microbiome datasets with paired virus and bacterial metagenomic sequence sets [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei]. Our approach built off of previous large-scale phage-bacteria microbiome network analyses by inferring interactions using metagenomic datasets, instead of using culture-based techniques [@Flores:2013hc]. This metagenomic interaction inference model was powered beyond previous models, e.g., [@Edwards:2015iz, @Roux:2016cc], by its inclusion of protein interaction data as well as both negative and positive interactions. Through this approach, we were able to provide a basic understanding of the network dynamics associated with healthy human phage and bacterial <!-- archaeal or "their microbial host"--> communities. By building and utilizing a microbiome-virome network, we found that different people, body sites, and anatomical locations not only support distinct microbiome membership and diversity [@Hannigan:2015fz; @Minot:2011ez; @Reyes:2010cwa; @Turnbaugh:2009ei; @Grice:2009eea; @Findley:2013jf; @Costello:2009im, @Consortium:2012iz], but also support communities with distinct communication structures and propensities toward community instability. Through an improved understanding of the healthy state of network structures across the human body, we empower future studies to begin investigating how these community structures are influenced by disease state and their overall impact on human health.

@@

## Cohort Curation and Sample Processing
We studied the differences in virome-microbiome interaction networks across healthy human bodies by leveraging previously published sequence sets containing purified virome samples paired with bacterial metagenomes from whole metagenomic shotgun sequences. Our study contained three datasets that explored the impact of diet on the healthy human gut virome [@Minot:2011ez], the impact of anatomical location on the healthy human skin virome [@Hannigan:2015fz], and the viromes of monozygotic twins and their mothers [@Reyes:2010cwa; @Turnbaugh:2009ei]. These datasets were selected because they were subjected to virus-like particle (VLP) purification. They employed a chloroform treatment <!--did they all do this? did they use the same purification treatment? --> to eliminate organismal DNA and thereby allowing for direct assessment of both the extracellular and fully-assembled intracellular virome **(Supplemental Figure \ref{SequenceStats} A-B)** [@Minot:2011ez, @Hannigan:2015fz, @Reyes:2010cwa; @Turnbaugh:2009ei].

Here, the bacterial and viral sequences from these studies were quality filtered and assembled into contigs. We generated clusters of related bacterial and phage contigs based on their k-mer frequencies and co-abundance patterns using the CONCOCT algorithm **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**. Contig lengths within clusters ranged between $10_{3}$ and $10_{5.5}$, confirming the majority of contigs represented genome fragments, rather than complete genomes  **(Supplemental Figure \ref{ContigStats} - \ref{ClusterStats})**. We interpreted these clusters as operational units of related bacteria and phage genomes, which we defined as operational genomic units (OGUs). These OGUs are conceptually similar to the operational taxonomic unit (OTU) and operational protein family (OPF) definitions used for grouping highly similar 16S rRNA gene and open reading frame sequences, respectively [@Schloss:2008ht].

@@

Four phage and bacterial genomic features were used to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq]. The resulting random forest model performed with an AUC of `r signif(genmodelper$ROC, digits = sigfig)`, a sensitivity of `r signif(genmodelper$Sens, digits = sigfig)`, and a specificity of `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**. The most important predictor in the model was amino acid similarity between genes (blastx), followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**. Protein family interactions were moderately important to the model, and CRISPRs were not informative <!-- how many of the bacteria had Cas proteins? At least 50%? How informative did you expect CRISPRs to be in your training set? And how does this relate to how informative we expect them to be across bacteria/archaea? Most work predicting phage-host interactions from metagenomic data have relied on CRISPR linkages, so we might want to contextualize why it was not informative, as those readers and community members will be wondering about this.--> **(Figure \ref{RocCurve} B)**. Approximately one third of the training set relationships yielded no score and were therefore unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.

@@

In addition to diet, we also observed an association between obesity and network structure **(Figure \ref{dietnetworks} C-D)**. The obesity-associated network demonstrated a higher degree centrality **(Figure \ref{dietnetworks} C)**, but less closeness centrality compared to the healthy controls **(Figure \ref{dietnetworks} D)**. These results suggest that the obesity network was overall less connected, having microbes further from all other microbes within the community.

## Individuality of Microbial Networks
Skin and gut community membership and diversity are highly personal, with people remaining more similar to themselves than to other people over time [@Grice:2009ee; @Hannigan:2015fz; @Minot:2013ih]. We therefore hypothesized that this personal conservation extended to microbiome network structure. We addressed this hypothesis by calculating the degree of dissimilarity between each subject's network, based on phage and bacteria abundance and centrality. We quantified phage and bacteria centrality within each sample graph using the weighted eigenvector centrality metric. This metric defines central phages as those that are highly abundant <!-- how was abundance determined? read mapping to contigs, "size" of OGUs, as in how many contigs belonged to an OGU group? a combination --> and infect many distinct bacteria which themselves are abundant and infected by many other phages. Similarly, bacterial centrality was defined as those bacteria that were both abundant and connected to numerous phages that were themselves connected to many bacteria. We then calculated the similarity of community networks using the weighted eigenvector centrality of all nodes between all samples. Samples with similar network structures were interpreted as having similar capacities for maintaining stability and transmitting genetic material.

@@

Skin microbiome network structure was strongly conserved within individuals (p-value < 0.001, **Figure \ref{intradiv} D**). This distribution was similar when separated by anatomical sites. Most sites were statistically significantly more conserved within individuals **(Supplemental Figure \ref{allskin})**.

## Association Between Environmental Stability and Network Structure Across the Human Skin Landscape
Extensive work has illustrated differences in XYZ <!-- diversity? community structure? --> in the healthy human skin microbiome between anatomical sites, including bacteria, viruses, and fungi [@Grice:2009ee; @Findley:2013jf; @Hannigan:2015fz]. These communities vary by degree of skin moisture, oil, and environmental exposure. As viruses are known to influence microbial XYZ <!-- above, maybe community structure and diversity --> , we hypothesized that microbe-virus network structure would be specific to anatomical sites, as well. To test this, we evaluated the changes in network structure between anatomical sites within the skin dataset.

The average centrality of each sample was quantified using the weighted eigenvector centrality metric. Intermittently moist skin sites (dynamic sites that fluctuate between being moist and dry) were significantly less connected than the more stable moist and sebaceous environments (p-value < 0.001, **Figure \ref{skinnetwork} A)**. Also, skin sites that were protected from the environment (occluded) were much more highly connected than those that were constantly exposed to the environment or only intermittently occluded (p-value < 0.001, **Figure \ref{skinnetwork} B)**.

To supplement this analysis, we compared the network signatures using the centrality dissimilarity approach described above. The dissimilarity between samples was a function of shared relationships, degree of centrality, and bacteria/phage abundance. When using this supplementary approach, we found that network structures significantly clustered by moisture, sebaceous, and intermittently moist status **(Figure \ref{skinnetwork} C,E)**. Occluded sites were significantly different from exposed and intermittently occluded sites, but there was no difference between exposed and intermittently occluded sites **(Figure \ref{skinnetwork} D,F)**. These findings provide further support that skin microbiome network structure differs significantly between skin sites.

# Discussion
<!-- Much of the discussion repeats the statements in the results, it felt a bit redundant. At times, the discussion even repeated itself (you'll see I deleted most of the last paragraph). There are probably ways to be less repetitive and more provocative or to bring in discussion of more recent work? Or to reduce it even more by removing redundancies. Alternatively, could remove the redundancies in the results...as in only state the results, and nt reference previous work. Then could discuss how findings relate to previous work only in the discussion, rather than the results...which does happen at times in current form.-->
Foundational microbiome work has provided a baseline understanding of the human microbiome by characterizing bacteria and viral diversity across the human body, as well as other environments [@Grice:2009eea; @Findley:2013jf; @Hannigan:2015fz; @Costello:2009im, @Consortium:2012iz; @Schloss:2005hz; @Minot:2011ez]. Here, we offer an initial understanding of how phage-bacteria networks differ throughout the human body, so as to provide a baseline for future studies of how and why microbiome networks differ in disease states. We developed and implemented a network-based analytical approach to evaluate the basic properties of the human microbiome through bacteria and phage relationships, instead of membership or diversity alone. This enabled the application of network theory to provide a new perspective on complex ecological communities. We utilized metrics of connectivity to model the extent to which communities of bacteria and phages interact (e.g., horizontal gene transfer, modulated bacterial gene expression, alterations in abundance <!--again, I would keep building on this, or just build on it once in the intro, but then be sure not to parenthetically distill it to only these three examples in the discussion here-->). <!-- what I deleted here is starting to be overstated throughout, no need to say it again here -->

Just as gut microbiome and virome composition and diversity are conserved in individuals [@Hannigan:2015fz; @Grice:2009eea; @Findley:2013jf; @Minot:2013ih], gut and skin microbiome network structures were conserved within individuals over time. Yet, gut network structure was not conserved among family members. These findings suggested that microbiome network properties, such as stability and the potential for horizontal gene transfer <!-- and lots of other things related to host and virus eco and evo processes -->, were personal and may be impacted by personal factors ranging from the body's immune system to environmental conditions, such as climate and diet <!-- though I'd argue that the immune system is a component/control on the environmental condition of the microbial communities -->. The ability of environmental conditions to shape gut and skin microbiome network structure was supported by our finding that diet and skin location were associated with altered network structures.

@@

In addition to diet, the skin microbiome network structure varied by skin environment. Network structure differed between environmentally exposed and occluded skin sites. The sites under greater environmental fluctuation and exposure, the exposed and intermittently exposed sites, were less connected and therefore were predicted to have a higher propensity for instability. Likewise, intermittently moist sites demonstrated less connectedness than the more stable moist and sebaceous sites. Together these data suggested body sites under greater degrees of fluctuation harbored less connected, potentially less stable microbiomes.

While these findings take us an important step closer to understanding the microbiome through interspecies relationships, there are caveats to the approach that should be noted. First, although our infection classification model is advantageous over existing models, we recognize that, like most classification models, ours is only as good as its training set. Large-scale experimental screens for phage and bacteria infectious interactions that report high-confidence negative interactions (i.e., no infection) will provide more robust model training and improved model performance. Just as we have improved on previous modeling efforts, we expect that new and creative scoring metrics will be integrated into this model to further improve performance. Second, although our analyses offer an informative proof of concept, this work was done retrospectively and relied on published research from as long as seven years ago. These archived datasets were limited by the technology and costs of the time, meaning the datasets were suboptimally powered for the statistical analysis we strive for today. Despite these limitations, we were able to present initial conclusions. Follow-up studies will validate our findings and inform the design and interpretation of future studies.

Our work takes a first step towards defining bacteria-virus interaction profiles as a characteristic of human-associated microbial communities. We provide a foundation for future studies to begin evaluating how microbiome network dynamics change in disease states, and how the information can be applied therapeutically.
